---
title: "Data Vis Assignment"
author: "mhorger"
date: "March 11, 2019"
output:
    html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE, 
                      eval = TRUE,
                      error = TRUE)
library(ggplot2)
```

# Data-vis assignment 

### Calcutt, S. E., Proctor, D., & Berman, S. M. (2019). Chimpanzees (Pan troglodytes) Are More Averse to Social Than Nonsocial Risk. Psychological Science, 30, 105–115

###Graph 1
```{r}
library(tidyverse)


chimps<- c("Dara", "Rita","Liza", "Azalea", "Barbi",  "Rowena", "Katie", "Fiona")

cond<- c("social","social","social", "social", "social", "social", "social", "social", "nonsocial","nonsocial","nonsocial","nonsocial","nonsocial","nonsocial","nonsocial","nonsocial")
FreqOfRisk<- c(59, 60, 51, 57, 49, 35, 25, 45, 0, 42, 5, 35, 3, 3, 0, 10)
graph1<-data.frame(chimps, FreqOfRisk, cond)

print(graph1)

#modified code
ggplot(data = graph1) +
  geom_line (mapping = aes(x = cond, 
                           y=FreqOfRisk,
                           group=chimps,
                           linetype=chimps) )

#Crump's code
ggplot(data = graph1, aes(x = cond, 
                          y=FreqOfRisk, 
                          group=chimps,
                          linetype=chimps))+
  geom_line()

```
### Graph 2

```{r}
AverageFreqSocialRisk<- c(5, 11, 33, 40, 46, 49, 57, 60)
AverageFreqNonSocial<- c(3, 0, 45, 5, 1, 25, 52, 58)
graph2<- data.frame(AverageFreqNonSocial, AverageFreqSocialRisk) 

ggplot(data = graph2)+
  geom_point (mapping = (aes (x = AverageFreqNonSocial, y = AverageFreqSocialRisk)))+
     geom_smooth(method='lm',mapping = (aes (x = AverageFreqNonSocial, y = AverageFreqSocialRisk)))
```

###Cao, J., Kleiman-Weiner, M., & Banaji, M. R. (2019). People Make the Same Bayesian Judgment They Criticize in Others. Psychological Science, 30, 20–31.

### Graph 3

```{r}
cases<- c("Doctor Study 1", "Doctor Study 1", "Doctor Study 1", "Butcher Study 2","Butcher Study 2", "Butcher Study 2", "Firefighter Study 2","Firefighter Study 2", "Firefighter Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2" )
options<- c("Woman more likely", "Equally likely", "Man more likely")
totalN<-c(199, 199, 199, 199, 199, 199, 200, 200, 200, 205, 205, 205)
responses<-c(0, 189, 10, 9, 130, 60, 0, 120, 80, 6, 124, 75 )

props<-c(responses/totalN)

graph3<- data.frame(cases, options, totalN, responses, props)

print(graph3)



ggplot(data = graph3) +
  geom_bar(mapping = aes(x = cases, color = options, position = "fill", stat(responses / totalN)))


```

colors are wrong 

### Graph 4

```{r}
evaluation<- c(1, 2, 3, 4, 5, 6, 7)

cases<- c("Doctor Study 1", "Doctor Study 1", "Doctor Study 1","Doctor Study 1", "Doctor Study 1", "Doctor Study 1", "Doctor Study 1", "Butcher Study 2","Butcher Study 2", "Butcher Study 2","Butcher Study 2","Butcher Study 2", "Butcher Study 2","Butcher Study 2", "Firefighter Study 2","Firefighter Study 2", "Firefighter Study 2","Firefighter Study 2","Firefighter Study 2", "Firefighter Study 2","Firefighter Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2", "Construction Worker Study 2" )

EvalScores<- c(35, 50, 42, 35, 20, 10, 7, 19, 30, 50, 50, 30, 15, 5, 25, 25, 40, 50, 40, 15, 5, 25, 30, 30, 35, 55, 35, 20 )

graph4<-data.frame(evaluation, cases, EvalScores)

print(graph4)


p <- ggplot(graph4, aes(x=cases, y=EvalScores)) + 
  geom_violin() 

p + stat_summary(fun.data=data_summary)

p + stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="crossbar", width=0.2 )
p
```

### Graph 5

```{r}
PersonX<- c("Equally likely", "Man more likely")
MoneyTransferred<-c("$0.00","$0.00", "$0.05","$0.05", "$0.10","$0.10", "$0.15","$0.15", "$0.20", "$0.20", "$0.25", "$0.25", "$0.30", "$0.30")
Frequ<- c(5, 110, 2, 20, 70, 30, 108, 35, 55, 5, 1, 1, 1, 1)
graph5<-data.frame(PersonX, MoneyTransferred, Frequ)

a<- ggplot(graph5, aes(x=PersonX, y=Frequ)) + 
  geom_violin() 

a + stat_summary(fun.data=data_summary)
 stat_summary(fun.data="mean_sdl", mult=1, 
                 geom="crossbar", width=0.2 )
a


b<- ggplot(data = graph5)+
geom_violin( mapping = aes(x=PersonX, y=Frequ, stats = )) 

```
### Graph 6 

What kind of data do I need to put in to do the log likelihood ratio?
```{r}

```


### Graph 7 
```{r}
library(tidyverse)
gender<- c("man", "woman")
judgements<-c("prior", "model posterior", "reported posterior")
values<- c(60, 90, 75, 30, 70, 76)
graph7<- data.frame(gender, judgements, values)


ggplot(graph7, aes(fill=judgements, y=values, x=gender)) + 
    geom_bar(position="dodge", stat="identity")+
  labs( y = "P(Target = Doctor") 







```


###Graph 8
```{r}
 





```

###Jefferies, L. N., & Lollo, V. D. (2019). Sudden Events Change Old Visual Objects Into New Ones: A Possible Role for Phasic Activation of Locus Coeruleus. Psychological Science, 30, 55–64.

### Graph 9


```{r}


condition<- c("No transient", "Transient")
place<-c("outside",  "outside","inside", "inside")
all_data<-c(305, 292, 300, 280, 295, 265, 290, 250)
vis_data<-c(305, 292, 300, 280)
aud_data<-c(295, 265, 290, 250)
group<- c("visual","visual", "auditory", "auditory")

visual<-data.frame(condition, place, vis_data)

auditory<-data.frame(condition, place, aud_data)

graph9<-data.frame(visual, auditory)

ggplot(graph9, aes(fill=place, y=vis_data, x=condition)) + 
    geom_bar(position="dodge", stat="identity")+
   ylim(230, 340)

ggplot(graph9, aes(fill=place, y=aud_data, x=condition)) + 
    geom_bar(position="dodge", stat="identity")+
   expand_limits(y=c(230, 340))

```
```{r}

```

### Jouravlev, O., Schwartz, R., Ayyash, D., Mineroff, Z., Gibson, E., & Fedorenko, E. (2019). Tracking Colisteners’ Knowledge States During Language Comprehension. Psychological Science, 30, 3–19.

###Graph 10

```{r}

```

### Kim, H., Domenico, S. I. D., & Connelly, B. S. (2019). Self–Other Agreement in Personality Reports: A Meta-Analytic Comparison of Self- and Informant-Report Means. Psychological Science, 30, 129–138.
### Graph 11


```{r}


OCEAN<- c("emotional stability", "extraversion", "openness/intellect", "agreeableness", "conscientiousness")
informants<-c("family","family","family","family","family", "friends","friends","friends","friends","friends", "colleagues","colleagues","colleagues","colleagues","colleagues", "strangers","strangers","strangers","strangers","strangers")
mean_diff<-c(-.10,-.10,.10,-.20,-.20, -.20, -.10, .20, -.05, -.20,-.20, -.05,.10, 0,  -.10, 0, .30, .70, .30, .30)

graph11<-data.frame(OCEAN, informants, mean_diff)
print(graph11)

ggplot(graph11, aes(x=OCEAN, y=mean_diff, group=informants)) +
  geom_line(aes(linetype=informants))+
  geom_point()+
expand_limits(y=c(-.80, .80))
```

