---
title: "Welcome to my TED talk"
subtitle: "about stuff"
author: "Melissa Noel Rose Horger"
date: "`r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

Class: inverse, center, middle
# Opening slide 

---

# Primary headers are indicated by 1 hashtag

## Secondary by two, etc

- we make bullet points by adding dashes 
- just like this 
- the code here doesn't include a class delineation so it goes to the default 


---

#slide number 3


add R code blocks 

```{r}
HP <- c("harry", "ron", "hermione")

```

If we type echo = FALSE, it will hide the code but show the answer 


```{r, echo=FALSE, fig.align='center'}

hist(rnorm(1000, 1, 100))

```

---
# How do we figure out the height of our figures?

```{r, fig.align= 'left', fig.height=5}

hist(rnorm(1000, 1, 100))

```

---

```{r, fig.align='right', out.width="200px"}
hist(rnorm(1000, 1, 100))
```

---

# Adding pictures 

```{r, echo=FALSE}


imag1_path <- "images/hpmeme.PNG"


knitr::include_graphics(imag1_path)

```

---
#Adding columns 

.pull-left[

-pros of Gryffindor. 
- cons of Gryffindor. 

]

.pull-right[

-pros of Ravenclaw.
- cons of Ravenclaw.
]



---

#Displaying one line at a time 


Type your first line 

--

DOUBLE DASH

--

Then the next line 

--

Et cetera 