---
title: "Dictionary"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#R Intrinsics


##List of functions
###General bits
help(topic) - command to search the help page

```{r}
help(sum)
```


?topic - shorter command to search the help page 

```{r}
?sum
?ls
?dir
```


ls() - list objects 
```{r}
ls(a)

```


dir()

list.files()

###Input and output
save()

load()

data()

library()

read.table()

read.csv()

scan()

print() = lists data, vectors, etc 

```{r}
print(a)
```

cat()

write.table()

###Data Creation
c() - combine. how we can add numbers to a vector 
```{r}
a<- c(a, "cat", "dog", "Pickles")
```


from:to (where from and to are replaced with numbers, e.g. 1:10)
```{r}
d<-c(21, 22, 23, 24, 25, 25, 8, 83, 7, 85, 3,2,6)
d<-c(3:7, 10)
1:10

```


seq() - generate a sequence of numbers 
```{r}
seq(20)
```


rep()

data.frame()

list()

matrix()

factor()

rbind()

cbind()

###Slicing and extracting data
####indexing vectors
x[n] nth element

x[-n] all but nth element

x[1:n] first n elements

x[-(1:n)] elements from n+1 to the end

x[c(1,4,2)] specific elements

x[“name”] elements named “name”

x[x>3] all elements greater than 3

x[x > 3 & x < 5] all elements between 3 and 5

x[x %in% c(“a”,“and”,“the”)] all elements in given set

####Indexing lists
x[n] list with elements n

x[[n]] nt h element of the list

x[[“name”]] element of the list named “name”

x$name id.

####Indexing matrices
x[i,j] element at row i, column j

x[i,] row i

x[,j] column j

x[,c(1,3)] columns 1 and 3

x[“name”,] row named “name”

####Indexing data frames (matrix indexing plus the following)
x[[“name”]] column named “name”

x$nameid.

###Variable conversion
as.data.frame(x)

as.numeric(x)

as.logical(x)

as.character(x)

###Variable information
is.na(x)

is.null(x)

is.data.frame(x)

is.numeric(x)

is.character(x)

length(x)

dim(x)

dimnames(x)

nrow(x)

ncol(x)

class()

attributes()

###Data selection and manipulation
which.max()

which.min()

which()

sort()

unique() - what are the unique/nonrepeated values in a string 
```{r}
b<-c(2, 2, 4, 5, 64, 8, 1, 6, 9)


unique(b)
```


table()

sample()

###Math
max() - lists the maximum number in a string 

```{r}
vector<- c(1, 2, 3, 4, 5)
max(vector)
```


min()- lists the maximum number in a string
```{r}
min(vector)
```

range() - lists the min and max of a string

```{r}
range(vector)
range(a)
```

sum() - adds the values 

```{r}
sum(vector)
sum(a)
```


mean() - average 
```{r}
mean(vector)
```

median() - middle value 

```{r}
median(vector)
```

var() - variance 

sd() - standard deviation 

cor()

round()

abs() - absolute value? 

###Matrices
t()

diag()

rowSums()

colSums()

rowMeans()

colMeans()

###Advanced Data processing
apply()

aggregate()

###Strings
paste()

strsplit() - used to split up character strings - it's called a string split 
```{r}
a<-"sdjflskdflksdjflkjs"
a[1]
strsplit(a, split ="s")
```


tolower()

toupper

###Plotting
hist()

plot()

###Distributions
rnorm()

runif()

###Programming
show that you can define a function
show that you can write a for loop
show that you can write a while loop
show that you can write an if else statement
Explain what return() does inside a function, show you can use it
Explain what break() does, show you can use it

##Installing some packages
Use the packages tab in R-studio to install these packages. You will need to be connected to the internet when you do this. If you are installing on your laptop, or on R-studio Cloud, then these packages will not need to installed again.

ggplot2
dplyr
shiny
data.table
reshape2
stringr