---
title: "Dictionary"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 5
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#R Intrinsics


##List of functions
###General bits
help(topic) - command to search the help page

```{r}
# help(sum)
```


?topic - shorter command to search the help page 

```{r}
# ?sum
# ?ls
# ?dir
```


ls() - list objects 
```{r}
# ls(a)

```


dir() - list the files in a directory or folder. 

```{r}
dir(path = ".", pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE,
           ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)
```


list.files() - different command, same output 
```{r}
list.files(path = ".", pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE,
           ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)

```


###Input and output
save() - writes an external representation of R objects  to a file 

save(..., list = character(),
     file = stop("'file' must be specified"),
     ascii = FALSE, version = NULL, envir = parent.frame(),
     compress = isTRUE(!ascii), compression_level,
     eval.promises = TRUE, precheck = TRUE)
   
```{r}
# save(mat, file = "mat.RData")

```


load()- Reload datasets written with the function save.
load(file, envir = parent.frame(), verbose = FALSE)


```{r}
# load(file = "mat.RData")
```



data()- Loads specified data sets, or list the available data sets.
data(..., list = character(), package = NULL, lib.loc = NULL,
     verbose = getOption("verbose"), envir = .GlobalEnv)
```{r}
# data(file = mat.RData)
# data
```
     

library() - load and attach add-on packages 

```{r}
library(ggplot2)
```


read.table() - Reads a file in table format and creates a data frame from it, with cases corresponding to lines and variables to fields in the file. 
```{r}

```


read.csv() - Reads a .csv file 

scan() - Read data into a vector or list from the console or file.
```{r}
(x <- scan("https://stats.idre.ucla.edu/stat/data/scan.txt", what = list(age = 0,
    name = "")))
```


print() = lists data, vectors, etc **can also be used as you are trying to de-bug 

```{r}
a<- c(1:5)
print(a)
```

cat() - outputs the objects, concatenating the representations 
cat is useful for producing output in user-defined functions. It converts its arguments to character vectors, concatenates them to a single character vector, appends the given sep = string(s) to each element and then outputs them.


write.table() - prints data into a file 

```{r}
write.table(b & d)
```


###Data Creation
c() - combine. how we can add numbers to a vector 
```{r}
# a<- c(a, "cat", "dog", "Pickles")
```


from:to (where from and to are replaced with numbers, e.g. 1:10)
```{r}
# d<-c(21, 22, 23, 24, 25, 25, 8, 83, 7, 85, 3,2,6)
# d<-c(3:7, 10)
# 1:10

```


seq() - generate a sequence of numbers 
```{r}
seq(20)
```


rep() - replicate the values in x
```{r}
x<-c (1, 2, 3, 4)

rep(x)

```



data.frame() - creates data frames, tightly coupled collections of variables which share many of the properties of matrices and of lists, used as the fundamental data structure by most of R's modeling software.

```{r}
x<-c (1, 2, 3, 4)
y<-c(a, b, c, d)
data.frame(x, y)
```


list() - prints the data in a variable 
```{r}
list(a)
```


matrix() - creates a matrix 
matrix(data = NA, nrow = 1, ncol = 1, byrow = FALSE,
       dimnames = NULL)
```{r}
hundreds <- c(matrix(data = 1:100, nrow = 10, ncol = 10))
```


factor() - The function factor is used to encode a vector as a factor (the terms ‘category’ and ‘enumerated type’ are also used for factors). If argument ordered is TRUE, the factor levels are assumed to be ordered. For compatibility with S there is also a function ordered.

rbind() - Take a sequence of vector, matrix or data-frame arguments and combine by rows, respectively. 

cbind() - Same as before but combines by columns 

###Slicing and extracting data
####indexing vectors
x[n] nth element 

```{r}
b[2]
```


x[-n] all but nth element

```{r}
b[-2]
```


x[1:n] first n elements
```{r}
x[1:6]
```


x[-(1:n)] elements from n+1 to the end
```{r}
a[-(1:3)]
```


x[c(1,4,2)] specific elements
```{r}
AllNumbers[c(4,5, 6)]
```


x[“name”] elements named “name”
```{r}

hey<-c("hi", "hey", "hello")
hey["hey"]
```


x[x>3] all elements greater than 3

```{r}
x[x>3]
```


x[x > 3 & x < 5] all elements between 3 and 5
- such as if you want items 5-10

x[x %in% c(“a”,“and”,“the”)] all elements in given set

####Indexing lists
x[n] list with elements n
```{r}
y[2]
```


x[[n]] nt h element of the list
```{r}
y[[5]]
```


x[[“name”]] element of the list named “name”

x$name id.

####Indexing matrices
x[i,j] element at row i, column j

```{r}
hundreds[1,2]
```


x[i,] row i

```{r}
hundreds[2,] 
```


x[,j] column j

x[,c(1,3)] columns 1 and 3

x[“name”,] row named “name”

####Indexing data frames (matrix indexing plus the following)
x[[“name”]] column named “name”

x$nameid.

###Variable conversion
as.data.frame(x) - Functions to check if an object is a data frame, or coerce it if possible.

as.numeric(x) - creates or coerces objects to numeric 

as.logical(x) - Create or test for objects of type "logical", and the basic logical constants.

as.character(x) - Create or test for objects of type "logical", and the basic logical constants.

###Variable information
is.na(x) - indicates which elements are missing 
```{r}
is.na(a)
```


is.null(x) - often returned by expressions and functions whose value is undefined.
```{r}
is.null(hey)
```


is.data.frame(x) - to check if an object is a data frame or coerse it if possible 

is.numeric(x) - general test of an object being interpretable as number 

is.character(x) - is an object interpretable as characters 

length(x) - number of items in a variable 

dim(x) - dimensions of a matrix 
```{r}
dim(hundreds)
```


dimnames(x) - retrieve or set the dimension of an object 
```{r}

```


nrow(x) - number of rows of an array 

ncol(x) - number of columns of an array 

class() - classify an object 

attributes() - list object's attributes 

###Data selection and manipulation
which.max()

which.min()

which()

sort()

unique() - what are the unique/nonrepeated values in a string 
```{r}
b<-c(2, 2, 4, 5, 64, 8, 1, 6, 9)


unique(b)
```


table() - create a table 


sample() - generate random sample 
```{r}
sample(1:100, 99)
```


###Math
max() - lists the maximum number in a string 

```{r}
vector<- c(1, 2, 3, 4, 5)
max(vector)
```


min()- lists the maximum number in a string
```{r}
min(vector)
```

range() - lists the min and max of a string

```{r}
range(vector)
range(a)
```

sum() - adds the values 

```{r}
sum(vector)
sum(a)
```


mean() - average 
```{r}
mean(vector)
```

median() - middle value 

```{r}
median(vector)
```

var() - variance 

sd() - standard deviation 

cor()

round() - round an integar 

abs() - absolute value? 

###Matrices
t()

diag()

rowSums()

colSums()

rowMeans()

colMeans()

###Advanced Data processing
apply()

aggregate()

###Strings
paste()

strsplit() - used to split up character strings - it's called a string split 
```{r}
a<-"sdjflskdflksdjflkjs"
a[1]
strsplit(a, split ="s")
```


tolower()

toupper

###Plotting
hist()

plot()

###Distributions
rnorm()

runif()

###Programming
show that you can define a function
show that you can write a for loop
show that you can write a while loop
show that you can write an if else statement
Explain what return() does inside a function, show you can use it
Explain what break() does, show you can use it

##Installing some packages
Use the packages tab in R-studio to install these packages. You will need to be connected to the internet when you do this. If you are installing on your laptop, or on R-studio Cloud, then these packages will not need to installed again.

ggplot2
dplyr
shiny
data.table
reshape2
stringr