---
title: "Using the lme function from the nlme package"
subtitle: "A brief tutorial"
author: "Melissa Horger"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, duke-blue, middlebury-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```



class: center, middle

# Let's consider the competition 

.pull-left[

##nlme. 
- Let's you specify variance-covariance structures for the residuals.
- Well suited for repeated measure or longitudinal designs. 

]

.pull-right[

##lme4.
- Allows you to fit outcomes whose distribution is not Gaussian and allows for crossed random effects.
- Stores data more effiently due to the use of sparse matrices.
- More suited for clustered data sets.
]


---


#What's inside? 

--

### A whole lot...

--
1. Sample data

--

2. Matrices

--

3. Statistics

--

4. Lattice framework


--

```{r, echo=FALSE, out.height="50%", out.width="50%", fig.align='center'}
imag1_path <- "images/obama.PNG"


knitr::include_graphics(imag1_path)
```

---

# The lme function within the nlme package

This generic function fits a linear mixed-effects model in the formulation described in Laird and Ware (1982) but allowing for nested random effects. The within-group errors are allowed to be correlated and/or have unequal variances.



--

This function allows us to answer research questions about how individuals change over time and variables that may influence development. 


--


**An example: Does the number of daily naps impact infant performance on a thing from 1 to 15 months?**


---

#Using nlme

## Begin by installing the nlme package

Found on the CRAN repository 


Website: https://svn.r-project.org/R-packages/trunk/nlme

```{r, eval=FALSE}
install.packages("nlme")
```


## Load the nlme package (and other useful packages)
```{r}
library(nlme)
library(ggplot2)
library(knitr)
library(dplyr)

```


---



#Create a data set

I created a data set based on the hypothesis that scores will improve with infant age and with less fragmented sleep(fewer naps). 

```{r}
Subs <- rep(c(seq(1:10)), 4)

Month <- c(rep(c(1), 10), rep(c(5), 10), rep(c(10), 10), rep(c(15), 10))

Naps <- c(rep(c(3), 10), 3, 3, 3, 3, 3, 2, 2, 2, 2, 2, 3, 3, 3 ,2,2,  2, 2,2,1, 1, 2, 2, 2, 2, 2, 1, 1, 1, 1, 1)
Napsfactor <- as.factor(Naps)

secondscores <- c(runif(10, 1, 10), runif(5, 5, 10),runif(5, 9, 17), runif(3, 10, 15), runif(5, 14, 22), runif(2, 20, 25), runif(5, 18, 23), runif(4,22, 27), runif(1, 27, 30) )

seconddataset <- data.frame(Subs, Month, Naps, secondscores, Napsfactor)


```


---

##Data should be in long format for our analysis 

```{r}
print(seconddataset)

#Napsfactor is a variable we will use when graphing our results. We do not use it for the analysis
```


---

#Summary statistics from our data set 
### One way to check our manipulation  

Remember our hypothesis was that there will be a positive relationship between age and scores and a negative relationship between naps and scores. 
```{r}
seconddemos <- seconddataset %>%
    group_by(Month, Naps) %>%
  summarise(mean_secondscore = mean(secondscores, na.rm=TRUE))


print(seconddemos)
```




---


#Run a conditional growth curve analysis 

```{r}
secondtutorial<-nlme::lme(secondscores ~ Month * Naps, random = ~ Month | Subs, data=seconddataset)
```

**lme(model, random, data)**


**model** - secondscores ~ Month * Naps
- We expect scores will be influenced by how old infants are (Month) and the number of Naps they take per day. 
- There may be an interaction between these predictors. 


**random** - random = ~ Month | Subs
- Random error comes from the fact that this is a within subjects design. The same infants are assessed at 1 month, 5 months, 10 months, and 15 months. 


**data** - data=seconddataset


---

#Put the results of the model in a nicer table 




```{r}
secondtut<- summary(secondtutorial)
secondtabl = secondtut$tTable 
print(secondtabl) 
```





---

#Graphing the results 

We want to look at individual subjects' trajectories of performance and average number of naps across time. 

```{r}
secondplot<-
ggplot2::ggplot(seconddataset, aes(x=Month, 
                                   y=secondscores, 
                                   color=Napsfactor, 
                                   shape = Napsfactor, 
                                   group=Subs), 
                xlim(1, 15), ylim(0, 25), xlab(Month) ) + 
  geom_point()+
  geom_line( color="grey")
secondplot<- secondplot + scale_x_continuous(name="Age (in months)",                                                limits=c(1, 15), 
                                             breaks = Month) +
  scale_y_continuous(name="Scores", limits=c(0, 30))


```


---

#Our results 

```{r}
print(secondplot)
```


---

#References
Curran, P. J., Obeidat, K., & Losardo, D. (2010). Twelve frequently asked questions about growth curve modeling. Journal of Cognition and Development : Official Journal of the Cognitive Development Society, 11(2), 121–136. doi:10.1080/15248371003699969

Magnusson, K. (2015). Uing R and lme/lmer to fit different two- and three- level longitudinal models. R Psychologist. Retrieved from https://rpsychologist.com/r-guide-longitudinal-lme-lmer#longitudinal-two-level-model

Maindonald, J. (2007). Chapter 10: Multi-level models and repeated measures. In J. Maindonald & J. Braun (Eds.), Data analysis and graphics using R: An example-based approach. Cambridge: Cambridge University Press.

Pinheiro, J. & Bates, D. (2019). Package 'nlme'. Retrieved from https://cran.r-project.org/web/packages/nlme/nlme.pdf 




