---
title: "nlme"
subtitle: "Fits and Compares Linear and Nonlinear Mixed Effects Models"
author: "Melissa Horger"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, duke-blue, middlebury-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```



class: center, middle

# Let's consider the competition 

.pull-left[

##nlme. 
- Let's you specify variance-covariance structures for the residuals.
- Well suited for repeated measure or longitudinal designs. 

]

.pull-right[

##lme4.
- Allows you to fit outcomes whose distribution is not Gaussian and allows for crossed random effects.
- Stores data more effiently due to the use of sparse matrices.
- More suited for clustered data sets.
]


---


#What's inside? 

--

### A whole lot...

--
1. Sample data

--

2. Lattice

--

3. Statistics

--

4. Matrices


--

```{r, echo=FALSE, out.height="50%", out.width="50%", fig.align='center'}
imag1_path <- "images/obama.PNG"


knitr::include_graphics(imag1_path)
```

---

# The nlme function within the nlme package

This generic function fits a nonlinear mixed-effects model in the formulation described in Lindstrom and Bates (1990) but allowing for nested random effects. The within-group errors are allowed to be correlated and/or have unequal variances.

### What did Lindstrom and Bates do for us? 

--

.pull-left[
```{r, echo=FALSE, out.height="80%", out.width="80%"}
imag2_path <- "images/equation1.PNG"


knitr::include_graphics(imag2_path)

```

]

.pull-right[
```{r, echo=FALSE, out.height="120%", out.width="120%"}
imag3_path <- "images/equation.PNG"


knitr::include_graphics(imag3_path)

```
]

--

A lot of math

---
# Some important considerations  

1. Need repeated measures from a single subject
  - May be longitudinal, but may not
  
2. Can account for correlations within individuals
  - within the random effects 
  
3. Uses maximum likelihood estimates 

---

#Using nlme

## Begin by installing the nlme package

Found on the CRAN repository 


Website: https://svn.r-project.org/R-packages/trunk/nlme

```{r, eval=FALSE}
install.packages("nlme")
```


## Load the package 
```{r, eval=FALSE}
library("nlme")
```

---

#Now we can view the specific nlme function



##The arguments for this function

nlme(model, data, fixed, random, groups, start, correlation, weights,
subset, method, na.action, naPattern, control, verbose)








