---
title: "Learning_R"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


# Practice Problems

## 1. Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+1
89392-8373 
4*3
122/12
```


## 2. Put numbers into variables, do simple math on the variables
```{r}
mel <-1
other<-23456789
a<-1
b<-20
c<-40

a+b

c-b

other/c 

(a+c)/b

```


## 3. Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r}
for(i in 1:100)
print (i)


seq(1,100)
```


## 4. Find the sum of all the integer numbers from 1 to 100.
```{r}
sum(i)

sum(1:100)

sum(seq(1,100))
```


## 5. Write a function to find the sum of all integers between any two values.
```{r}
sum(1:100)
sum(5:500)
sum(2:4)
```


## 6. List all of the odd numbers from 1 to 100.
```{r}
EvenNumbers<-c()
for(i in 1:100){
if (i%%2 ==0){
EvenNumbers<-c(EvenNumbers,i)
}
}

EvenNumbers
```



## 7. List all of the prime numbers from 1 to 1000.
```{r}
AllNumbers<-c()
for(i in 1:1000){
  AllNumbers<-c(AllNumbers,i)
}
PrimeNumbers<-c()
for (j in AllNumbers){
  if(j %/% 2 || j%/% 3)
    PrimeNumbers<-c(AllNumbers,j)
}

PrimeNumbers
```

```{r}
sieve <- function(n)
{
   n <- as.integer(n)
   if(n > 1e6) stop("n too large")
   primes <- rep(TRUE, n)
   primes[1] <- FALSE
   last.prime <- 2L
   for(i in last.prime:floor(sqrt(n)))
   {
      primes[seq.int(2L*last.prime, n, last.prime)] <- FALSE
      last.prime <- last.prime + min(which(primes[(last.prime+1):n]))
   }
   which(primes)
}

 sieve(1000)
```

## 8.Generate 100 random numbers
```{r}
sample(1:1000000, 100, replace=FALSE)
sample(1:10000, 100, replace=TRUE)
```


## 9. Generate 100 random numbers within a specific range
```{r}
sample(20:400, 100, replace=FALSE)
```


## 10.Write your own functions to give descriptive statistics for a vector variable storing multiple numbers. Write functions for the following without using R intrinsics: mean, mode, median, range, standard deviation

```{r}
a<-c(2,4,6,8,10,12,14,16)
b<-c(2,4,4,6)

(sum(a))/length(a)

b[duplicated(b)]


b[length(b)/2]
a[length(a)/2]



(a[length(a)])-(a[1])



average<-c(sum(a))/length(a)
average
d<-c((a-average)^2)
d 
sqrt((sum(d))/length(d))

sd(a)

```
## 11. Count the number of characters in a string variable


```{r}
Captain<- c("pickles", "is", "the", "best", "cat")
Class<-c( "this class is comprised of learning r")

nchar(Class)
nchar (Captain)

```

## 12. Count the number of words in a string variable


```{r}
sapply(strsplit(Class, " "), length)
sapply(strsplit(Captain, " "), length)

```


## 13. Count the number of sentences in a string variable

```{r}
paragraph<-c("A paragraph is a brief piece of writing that's around seven to ten sentences long. It has a topic sentence and supporting sentences that all relate closely to the topic sentence. The paragraph form refers to its overall structure, which is a group of sentences focusing on a single topic.")

sapply(strsplit(paragraph, "[.]"), length)

strsplit(paragraph, "[.]")
```
This answer is much too high. Unsure where this problem is coming from 

## 14. Count the number of times a specific character occurs in a string variable

```{r}
sapply(strsplit(paragraph, "paragraph"), length)
sapply(strsplit(paragraph, "n"), length)
```


## 15. Do a logical test to see if one word is found within the text of another string variable.

```{r}
a==b
a!=Captain
word<-"fart" 
word==Captain
word==paragraph
strsplit(paragraph,split=" ")
strsplit(Captain,split=" ")
Captain==paragraph

Captain%in%paragraph 

```
*not currently reading the all of the text within paragraph. should it be able to do that? 


## 16. Put the current computer time in milliseconds into a variable

```{r}
print(as.numeric(Sys.time())*1000, digits=10)
time<- c(print(as.numeric(Sys.time())*1000, digits=10))
```
I can't read milliseconds so let's try it normally... 

```{r}
time<- c(time, print(as.numeric(Sys.time())*100, digits=10))
Sys.time()

time<-c(Sys.time())
print (time)

```


## 17. Measure how long a piece of code takes to run by measuring the time before the code is run, and after the code is run, and taking the difference to find the total time

```{r}
time1<- print(as.numeric(Sys.time())*1000, digits=10)

seq(10000000)

time2 <- print(as.numeric(Sys.time())*1000, digits=10)

time2-time1 
```

## 18. Read a .txt file or .csv file into a variable

```{r}

breweries <- read.csv("Breweries .csv")
HistoryOfBeer <- read.delim("TestFile.txt")

```


## 19. Output the contents of a variable to a .txt file

```{r}

write(AllNumbers, file = "data.txt",
      ncolumns = 1,
      append = FALSE, sep = " ")

```


## 20. Create a variable that stores a 20x20 matrix of random numbers
```{r}
?matrix

mat <- matrix(data = rexp(200, rate = 10), nrow = 20, ncol = 20)

matr<-matrix(data = rnorm(4000, mean = 25, sd = 5), nrow = 20, ncol = 20)

?rexp
?Normal

```

