---
title: "Learning_R"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


# Problems 1-10 

## 1. Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+1
89392-8373 
4*3
122/12
```


## 2. Put numbers into variables, do simple math on the variables
```{r}
mel <-1
other<-23456789
a<-1
b<-20
c<-40

a+b

c-b

other/c 

(a+c)/b

```


## 3. Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r}
for(i in 1:100)
print (i)


seq(1,100)
```


## 4. Find the sum of all the integer numbers from 1 to 100.
```{r}
sum(i)

sum(1:100)

sum(seq(1,100))
```


## 5. Write a function to find the sum of all integers between any two values.
```{r}
sum(1:100)
sum(5:500)
sum(2:4)
```


## 6. List all of the odd numbers from 1 to 100.
```{r}
EvenNumbers<-c()
for(i in 1:100){
if (i%%2 ==0){
EvenNumbers<-c(EvenNumbers,i)
}
}

EvenNumbers
```



## 7. List all of the prime numbers from 1 to 1000.
```{r}
AllNumbers<-c()
for(i in 1:1000){
  AllNumbers<-c(AllNumbers,i)
}
PrimeNumbers<-c()
for (j in AllNumbers){
  if(j %/% 2 || j%/% 3)
    PrimeNumbers<-c(AllNumbers,j)
}

PrimeNumbers
```

```{r}
sieve <- function(n)
{
   n <- as.integer(n)
   if(n > 1e6) stop("n too large")
   primes <- rep(TRUE, n)
   primes[1] <- FALSE
   last.prime <- 2L
   for(i in last.prime:floor(sqrt(n)))
   {
      primes[seq.int(2L*last.prime, n, last.prime)] <- FALSE
      last.prime <- last.prime + min(which(primes[(last.prime+1):n]))
   }
   which(primes)
}

 sieve(1000)
```

## 8.Generate 100 random numbers
```{r}
sample(1:1000000, 100, replace=FALSE)
sample(1:10000, 100, replace=TRUE)
```


## 9. Generate 100 random numbers within a specific range
```{r}
sample(20:400, 100, replace=FALSE)
```


## 10.Write your own functions to give descriptive statistics for a vector variable storing multiple numbers. Write functions for the following without using R intrinsics: mean, mode, median, range, standard deviation

```{r}
a<-c(2,4,6,8,10,12,14,16)
b<-c(2,4,4,6)

(sum(a))/length(a)

b[duplicated(b)]


b[length(b)/2]
a[length(a)/2]



(a[length(a)])-(a[1])



average<-c(sum(a))/length(a)
average
d<-c((a-average)^2)
d 
sqrt((sum(d))/length(d))

sd(a)

```
## 11. Count the number of characters in a string variable

## 12. Count the number of words in a string variable

## 13. Count the number of sentences in a string variable

## 14. Count the number of times a specific character occurs in a string variable

## 15. Do a logical test to see if one word is found within the text of another string variable.

## 16. Put the current computer time in milliseconds into a variable

## 17. Measure how long a piece of code takes to run by measuring the time before the code is run, and after the code is run, and taking the difference to find the total time

## 18. Read a .txt file or .csv file into a variable

## 19. Output the contents of a variable to a .txt file

## 20. Create a variable that stores a 20x20 matrix of random numbers

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the sam place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation =1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```

When you knit this R Markdown document, you will see that the histogram is printed to the page, along with the R code. This document can be set up to hide the R code in the webpage, just delete the comment (hashtag), from the cold folding option in the yaml header up top. For purposes of letting yourself see the code, and me see the code, best to keep it the way that it is. You learn all of these things and more can be customized in each R code block.

# The big idea

Use this lab journal to record what you do in R. This way I will be able to see what you are doing and help you along the way. You will also be creating a repository of all the things you do. You can make posts about everything. Learning specific things in R (project unrelated), and doing things for the project that we will discuss at the beginning of the Fall semester. You can get started now by fiddling around with googling things, and trying stuff out in R. I've placed some helpful starting links in the links page on this website

# What can you do right now by yourself?

It's hard to learn programming when you don't have specific problems that you are trying to solve. Everything just seems abstract.

I wrote an [introductory programming book that introduces R](https://crumplab.github.io/programmingforpsych/), and gives some [concrete problems for you to solve](https://crumplab.github.io/programmingforpsych/programming-challenges-i-learning-the-fundamentals.html). 

To get the hang of journaling and solving the problems to learn programming, my suggestion is that you use this .Rmd file to solve the problems. It would look like this:

# Problem 1

Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+2
2*5
5/3
(1+6+4)/5

```

# Problem 2

Put numbers into variables, do simple math on the variables

```{r}
a<-1
b<-2
a+b

d<-c(1,2,3)
e<-c(5,6,7)
d+e
d*e
d/e

```

# Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r}
# for loop solution
# i becomes the number 1 to 100 at each step of the loop


a <- length(100) # make empty variable, set length to 100
for (i in 1:100){
  a[i] <-i #assigns the number in i, to the ith index of a
}

print(a)

# for loop solution #2

a<-c() #create empty variable using combine command
for (i in 1:100){
  a<-c(a,i) # keeps combining a with itself and the new number in i
}
print(a)

# seq solution
a <- seq(1,100,1) # look up help for seq using ?seq() in console
print(a)

```

# Replace this with problem 4

And keep going. Try to solve the problems with different scripts that provide the same solution. Good luck, Happy coding.
