---
title: "Statistics Problems"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 6
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message = FALSE, eval = FALSE)
```

##Statistics problems 

###Null distribution of mean differences
If you take two random samples from the same distribution they can be different because of random sampling. This question asks you to construct a null-distribution using simulation.

Assume your samples come from a normal distribution with mean =0, sd =1. (use rnorm)
Assume sample size for each sample is n =10 (e.g., rnorm(10,0,1))

```{r}
counter<-c(0)

for(i in 10000)

sample1<- rnorm(10000,0,1)

sample2<- rnorm(10000,0,1)

diff<- sample1 - sample2
```


Create the null-distribution: Run a simulation 10,000 times that finds the difference between means of sample A and B. 

```{r}
counter<-c(0)

for(i in 1:10000)
  sample1[i]<- rnorm(10000,0,1)
  sample2[i]<- rnorm(10000,0,1)
  diff[i]<- mean(sample1)-mean(sample2)

hist(diff)
```



You should have a vector of 10,000 difference scores
Assume an alpha criteria of p<.05 for a directional test. What is the critical value for a positive mean difference? In other words, how large must a positive mean difference be in order to occur less than 5% of the time under the null?
Assume an alpha criteria of p<.05 for a non-directional test. What is the critical value for the absolute value of the mean difference. In other words, how large must the absolute value of the mean difference be in order to occur less than 5% of the time under the null?