---
title: "nlme Tutorial"
author: "mhorger"
date: "May 4, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# nlme is a package for fitting and comparing linear and nonlinear mixed effects models
It let's you specify variance-covariance structures for the residuals.
It is well suited for repeated measure or longitudinal designs. 

###Similar packages
One similar package is lme4. 
It allows you to fit outcomes whose distribution is not Gaussian and allows for crossed random effects.
It stores data more effiently due to the use of sparse matrices.
It is more suited for clustered data sets.

##What's included? 
nlme includes sample data, statistics, matrices, and a lattice framework. 



#Using the nlme package 

## Begin by installing the nlme package

Found on the CRAN repository 


Website: https://svn.r-project.org/R-packages/trunk/nlme

```{r, eval=FALSE}
install.packages("nlme")
```


## Load the package 
```{r, eval=FALSE}
library("nlme")
```


## Try the lme function
This generic function fits a linear mixed-effects model in the formulation described in Laird and Ware (1982) but allowing for nested random effects. The within-group errors are allowed to be correlated and/or have unequal variances.

### Some important considerations  

1. Need repeated measures from a single subject The data may be longitudinal, but they also may not.
  
2. Can account for correlations within individuals within the random effects 
  
3. Uses maximum likelihood estimates 

###The arguments for this function

lme(model, data, fixed, random, groups, start, correlation, weights,
subset, method, na.action, naPattern, control, verbose)


```{r}

```

